### 2019-05-31 BiometricsPaper/bp_relatedWork.tex

%!TEX root = bioPrediction_main.tex
\section{Related Work}
We review two categories of related work for this study. First, we review work related to the productivity-related indicators that we considered in our study, including how they are detected and prevented in several risk-inducing scenarios. Second, we review work related to biometrics, including how they have been studied in the office environment or applied to assess cognitive states and processes.

\subsection{Productivity-related Indicators}

Previous studies have sought to measure, detect, and predict productivity indicators.

\subsubsection{Stress}
Much previous work relates to identifying and mitigating stress in an office environment.
Previous studies have measured stress by taking one of two possible approaches.
The first approach is to measure plasma catecholamine and cortisol as stress biomarkers~\cite{piazza10}.
This approach is impractical for use over prolonged periods of time, as in our eight-week study.
Further, this approach is imprecise because of the delay from the stress stimulation to the stress response, which 
may take from minutes to hours~\cite{Chandola10,Hellhammer09}.

The second approach, which we have chosen for our study, is to measure autonomic nervous system (ANS) activity by analyzing biometric signals of the human body, such as blood pressure, heartbeat, and temperature~\cite{kataoka00,Eekelen04,valentini10}.
In particular, changes in heart rate variability are associated with cognitive and emotional stress~\cite{mcduff16,dishman2000stress}.
This second approach has been used successfully by several past studies~\cite{Force96,gal07,montano09}.

Stress in the workplace and its effects on service providers has been analyzed in call centers~\cite{Hernandez11} and in the context of the perceived imbalance between resources and demands~\cite{cherniss80}. These studies considered several factors such as personality traits, career-related goals and attitudes, and life outside of work, and these factors were correlated with stress levels and burnout.

Evans and Johnson~\cite{evans00} investigated the correlation between noise in the workplace and stress levels.
They found that workers exposed to open-office noise showed aftereffects that indicate motivational deficits.
The population for their experiment comprised 40 female clerical workers, who were randomly assigned to a control condition or to three-hour low-intensity noise room designed to simulate typical open-office noise levels.

Hovsepian et al.~\cite{Hovsepian15} have worked to obtain a standard for continuous stress assessment. They used sensors to conduct a seven-day lab study with 26 participants, as well as a field study with 20 participants.

\subsubsection{Awakeness}

Sleepiness (lack of awakeness) and its associated risk of serious injury to passengers has been studied in the context of automobile 
accidents~\cite{connor02,Nordbakke07}.
These studies show a strong association between the level of acute driver sleepiness and the risk of injury crash.
Connor et al.~\cite{connor02} conducted a population-based case study using the Stanford sleepiness scale, which is similar to a seven-point Likert scale and describes seven different levels of sleepiness from ``Could not stay awake, sleep onset was imminent'' (1) to ``Felt active, wide awake'' (7).
Nordbakke and Sagberg~\cite{Nordbakke07} show that drivers are well aware of  various factors influencing the risk of falling asleep while driving. Drivers also have good knowledge of the most effective measures to prevent falling asleep at the wheel.
However, most of drivers continue driving even when recognizing sleepiness signals, due to the desire to arrive at a reasonable time, the length of the drive, or pre-planed commitments.

%Bader~\cite{Bader95} proposed an approach towards monitoring and estimating a person's awakeness by leveraging on a stationary pressure sensor to track the person's body movements and a correlator to produce a change signal which is compared with a preset wakefulness threshold in a threshold detector circuit.

%A wearable system for mood assessment considering smartphone features and data from mobile ECGs. Exler, Schankin, 
%Looked at energetic-arousal (tired-awake)
%This paper focuses on valence, and barely mentions energetic arousal 
%so I will leave it out

\subsubsection{Focus}
Focus refers to the allocation of limited cognitive processing resources~\cite{Anderson04}.
Mark et al.~\cite{mark2014bored} studied engagement in workplace activities by analyzing the digital activity of 32 information workers in situ for 5 days to understand how attentional states change with context. They found that boredom is highest in the early afternoon and focus peaks in the middle of the afternoon. They also found that doing work that requires focus correlates with stress, while rote work correlates with happiness.

Interruptions in the office are a common barrier keeping workers from sustaining focus on their work related activities, particularly when the interruptions occur at inopportune moments.
Such interruptions may include emails, alerts, or interactions with co-workers\cite{gonzalez2004constant,chong2006interruptions,shamsi07}. Interruptions in inopportune times can have negative effects that range from higher error rate and lower overall performance
to an increase in stress and frustration~\cite{bailey2001effects,czerwinski2000instant,mark2008cost}.
External interruptions may cause workers to enter a ``chain of distraction''~\cite{shamsi07}.
This chain is composed by stages of preparation, diversion, resumption and recovery that result in time away from an ongoing task.

Other studied constructs that relate to focus in the workplace include cognitive absorption, cognitive engagement, flow, and mindfulness.
Cognitive absorption describes periods of time in which a person experiences total immersion in an activity.
This state is also accompanied by a sense of deep enjoyment, a feeling of control, curiosity, and not realizing the passing of time.
It has been associated with ease of use and perceived usefulness of information technology~\cite{agarwal00}.
Cognitive engagement is described~\cite{webster97} as a period of strong focus in an activity without the feeling of a sense of control of the situation.
Flow~\cite{Csikszentmihalyi90}, and mindfulness~\cite{Weick06,dane11} are psychological states that describe periods of prolonged attention and total immersion in an activity. Flow occurs when a person is focused on an activity that requires high challenge and high use of the person's skills, whereas mindfulness is characterized by being aware of fine detail, affording the capacity to discover and manage unexpected events.

\subsection{Biometrics}
Our study investigates the prediction of multiple productivity-related indicators (stress, focus, and awakeness) using two different types of measurements, biometric signals and computer interaction, over eight weeks in a real-life office setting.
Existing work~\cite{sano2013stress,healey2005detecting,wijsman2011towards,zuger2015interruptibility,goyal2017intelligent,Parnin2011,Nakagawa2014,Radevski2015} analyzes a broad array of biometric signals and correlates them with individual's cognitive states and processes.
For example, Zuger et al.~\cite{zuger18} used biometrics to sense interruptibility in the office~\cite{zuger18}.
%and found correlations between biometric signals and predicting code quality~\cite{Muller16}, to assess task difficulty in software development~\cite{fritz2014using}, to assess developer's mental load and perceived difficulty while working on small code snippets, and
Biometric signals have also been studied in the context of technology users. For example, Parnin~\cite{Parnin2011} analyzes electromyography to measure sub-vocal utterances, and how these might be correlated with the programmer's perceived difficulty of programming tasks. Similarly, biometrics have been used to measure code difficulty by using biometric sensors~\cite{fritz2014using}
and using Near Infrared Spectroscopy to measure developer's cerebral blood flow~\cite{Nakagawa2014}.

Eye tracking technology~\cite{Bednarik2006,Crosby1990,Rodeghero2014} and brain activity~\cite{Ikutani2014,Siegmund2014} 
have been used in previous studies to analyze different tasks in an office environment.
Eye tracking has been used to analyze memory load and processing load by inspecting task-evoked pupillary response and pupil size~\cite{beatty82}. Similar studies have shown high correlation between pupil size and mental workload of subtasks~\cite{beatty82} and cognitive load~\cite{Klingner10}.
Brain activity has been associated with different mental states~\cite{Berger29} by analyzing specific frequency bands (alpha, beta, gamma, delta, and theta) using electroencephalography (EEG). The increase or decrease of some of these frequencies is correlated with
attentional demand and working memory load~\cite{Smith05,sterman93}.
In contrast to studies that use eye tracking or EEG, we focused on a less invasive technology that can be applied in a real world scenario.