### 2019-05-31 bp_conclusion.tex
merging the paper folder and the project folder

%!TEX root = bioPrediction_main.tex
\section{Conclusions} 
Stress, awakeness and focus at work are highly relevant aspects when it comes to productivity and well-being at the workplace. In this paper, we presented the results of a study with 14 professional knowledge workers in their workplace over an eight week period to examine the ability of biometrics for predicting these producitivity-related aspects. The longitude and in situ placement of the study---two key components of the study---as well as the breadth of human aspects examined, support and extend previous work. Based on twice collected daily survey responses and continually collected biometrics data we were able to create a model that was able to predict awakeness, stress, and focus, outperforming the baseline by as much as 84.6\% in the case of awakeness. The analysis further shows that there is a continuous increase in performance with the number of training instances, suggesting that there is more potential to improve the model's performance with further data points. Interestingly, while aspects such as stress have previously been predominantly linked to biometrics, our comparison with a model build from computer interaction features shows that even these types of features might bear big potential in predicting some of these productivity-related human aspects in the workplace.

The results also open up new opportunities to help increase knowledge workers' productivity and well-being, ranging from instantaneously taking action to prevent potentially risky situations and prevent accidents due to a lack of focus or awakeness, all the way to recommending interventions to reduce stress if it becomes more chronic.

% \todo{as you can see, the conclusion is way too long and has some cut and paste repetition}

% Stress, awakeness and focus at work are highly relevant aspects when it comes to productivity and well-being at the workplace. In this paper, we presented the results of a study with 14 professional knowledge workers in their workplace over an eight week period to examine the ability of biometrics for predicting these producitivity-related aspects. The longitude and in situ placement of the study as well as the breadth of human aspects examined, support and extend previous work. Based on twice collected daily survey responses and continually collected biometrics data we were able to create a model that was able to predict , outperforming the baseline by as much as 84.6\%, in the case of awakeness. We also show that there is a positive correlation between number of training instances and model performance--a 114\% improvement from training with one sample to training with 33 samples--suggesting that with a larger amount of training data prediction would continue to improve. Due to this, while current user-specific models have higher performance than across-user models, there is hope that with enough data even across-user models could become accurate.

% Our analysis shows that we are able to \todo{fill this}

% Our study with 14 professional knowledge workers in their workplaces over an eight week period has shown that we can use biometrics to predict these producitivity-related aspects of an individual knowledge worker. The longitude and in situ placement of the study as well as the set of human aspects examined support and extend previous work. In addition, our analysis

% as well as computer interaction features to predict these aspects of an individual knowledge worker
% has shown that we can use biometrics as well as computer interaction features to predict these aspects of an individual knowledge worker over a long period of time in situ
% The results also show that the general ... \todo{fill this}

% In this study we analyze biometric signals of office workers to predict the productivity-related human aspects of stress, focus, and awakeness. A key component of this study was its length, an eight week period, longer than previous studies by a factor of four~\cite{zuger18,Muller16}, and its realism , studying professionals in situ instead of students. We collected information from 14 participants from different departments within the company, covering different ages ranges, genders, and work experience. By learning from twice daily survey responses and continually collected biometrics data we were able to create a model that was able to predict human aspect values, outperforming the baseline by as much as 84.6\%, in the case of awakeness. We also show that there is a positive correlation between number of training instances and model performance--a 114\% improvement from training with one sample to training with 33 samples--suggesting that with a larger amount of training data prediction would continue to improve. Due to this, while current user-specific models have higher performance than across-user models, there is hope that with enough data even across-user models could become accurate.

% From these results emerge a series of opportunities to help increase awakeness, focus, and reduce stress in office workers. In the near future, we will therefore be able to detect a decrease in productivity-related aspects following our approach, and be able to take action to prevent potentially risky situations.

%\begin{acks}
%The acknowledgments section will be included for the camera-ready version of the paper.
%\end{acks}