### 2019-09-20 bp_results_1.tex
Restructured section levels and removed detail on classifiers.

%!TEX root = bioPrediction_main.tex
%\input{bp_results_1}
%%\input{bp_results_2}

%\section{Analysis and Results}

\begin{table*}[h]
  \centering
  \includegraphics[width=1.0\textwidth]{rq1performance.pdf}
  \caption{Results of predictions using the individual models. The distribution columns show a bar chart of the response distribution (negative/positive) for each of the three variables. The baseline row represents the averaged results of our baesline classifier. The general row shows the averaged results of our models trained on all participants.}\label{tab:accuracy}%\todo{revise caption}}
  \vspace*{-4mm}
\end{table*}

%% To evaluate the efficacy of continuously predicting a knowledge worker's stress, focus, and awakeness in the workplace, we trained and tested machine learning classifiers using a leave-one-out cross validation. For prediction, we used the features extracted from the collected data as the input data, and binarized the participants' self-reported responses on stress, focus, and awakeness into two classes each (e.g., 'stressed' and 'not stressed') to use as output measures.

\section{Observed Trends Over Time in Stress Levels}

To gain insights into how knowledge workers experience stress over an
extended period of work, we examined the end of day survey responses
collected from each participant to see if any identifiable trends
emerged. We used data from 13 of the 14 participants - we excluded one participant
from this analysis as they experienced atypical stress
levels in the latter half of the study due to factors outside of our
control.

\subsection{Baseline Stress Levels}
Common amongst all participants was a trend to select one stress rating
far more frequently than any other. We will refer to this value as the
participant's baseline stress level. All but one participant reported
their perceived stress level for the day as their baseline stress
level more than 50\% of the time. In total, the baseline values made
up 65\% of the reported values collected from
participants. Interestingly, while participants sometimes saw periods
of sustained increases in stress, lasting as many as 6 consecutive
workdays in the most extreme case, participants would always return to
their baseline stress level at some point.

The baseline stress level varied significantly between participants. 7
participants (54\%) reported feeling average stress levels most frequently
(rating 3 on our scale), while 5 (38\%) reported feeling little stress
(rating 2) and 1 (8\%) reported feeling no stress at all (rating 1).

\subsection{Stressful Days Tend to Cluster}
Accounting for the variance between participants perceived stress
baselines, we consider a stressful day to be one that represents a
deviation of 1 or more stress levels above the participant's
baseline. Of the 93 stressful days we observed in total, we found that
39 (41\%) of these days occurred in groupings of two or more
consecutive stressful workdays. The most common size of these groups
was two workdays, while the largest group we observed was six
workdays.  The day after a stressful day is much more likely to be a
stressful day as compared to any other day with a 0.55 average increase
over baseline, compared to 0.02 average increase over baseline.

\subsection{Extreme Changes in Stress Levels are Rare}
After accounting for each participant's perceived stress baseline, we
examined the frequency of deviations from the baseline. We found that
participants were far more likely to report a stress level that was
within 1 point of their baseline, than to report a stress level 2 or
more points away. These extreme deviations represented only 15\% of
all reported values which differed from the participants baseline. As
well, the majority (78\%) of these deviations came from just two
participants. This suggests that some people may be less resilient to
the stress of the workplace than others. For most participants,
extremely stressful days were few and far between.

\subsection{Explaining Stress Fluctuations}
We created a linear mixed model to attempt to explain some of the
stress that our participants were experiencing. We experimented with
multiple independent variables, including those derived from computer
interaction data such as total time spent actively on the computer,
percentage of time idle, and amount of time spent on non work-related
web browsing. We also looked at the day of the week and proximity to
beginning or end of month as possible explanatory variables. However,
ultimately we were unable to find any variable which showed a
significant correlation with our participants perceived stress
levels. This points to the need for additional instrumentation if we
are to succesfully understand and predict stress in the workplace.

%% \subsubsection{Tai Chi Intervention}
%% We linked participants tai chi session attendance data with their self-reported percieved stress levels. Building a linear mixed model with tai chi attendance as a random effect, we found that tai chi attendance contributed a small amount to decreased stress in the work week immediately following the tai chi session (slope of -0.188, p < 0.05). However, we did not find a connection between attendance and stress on the day of the session.

\section{Predicting Stress in the Moment}
% random forest best
\label{secOverallAccuracy}

To investigate whether stress, focus and awakeness can be predicted in
the moment based on biometric measures.
we investigated classifiers trained
for each individual and across all participants. We report on the effectiveness of
these classifiers and the features that are important in prediction of stress, focus
and awakeness.

\subsection{Selecting a Classifier Algorithm}

There are many different algorithms that can be used to build a classifier. To select an
algorithm, we compared multiple classifiers using the popular machine learning library scikit-learn~\cite{pedregosa11} and performed a grid search analysis to determine the optimal hyperparameters for each classifier. Our analysis showed that random forest outperforms all other classifiers, including Na\"ive Bayes, decision trees, support vector machine, and a multilayer perceptron neural network. The optimal values for random forest and the three output measures are listed in Table \ref{tab:hyperparams}. For the remainder of this paper, we refer to random forest classifiers trained with these hyperparameters.%\\[-0.1cm]
% for stress they were (minimum samples for split = 4, # of estimators = 100, max features = 0.5, k=300), for focus they were (minimum samples for split = 4, # of estimators = 50, max features = 0.5, k=200), and for awakeness they were (minimum samples for split = 4, # of estimators = 100, max features = 0.25, k=800)
\begin{table}[h]
	\begin{centering}
	\small\addtolength{\tabcolsep}{-1pt}
    \begin{tabular}{llll}
      \hline
      Variable & K & \# Estimators & Minimum Samples Split \\
      \hline
      Stress & 300 & 100 & 4\\
      Focus & 200 & 50 & 4\\
      Awakeness & 800 & 100 & 4\\
      \hline
    \end{tabular}
    \caption{The hyperparameters selected by grid search analysis to tune our random forest models. K refers to the number of features selected.}    \label{tab:hyperparams}
    \end{centering}
\end{table}

\subsection{Individual Classifiers}
%\noindent\textit{Results of Individual Classifiers}\\
Since peoples' experience of stress, focus, and awakeness (as well as
their physiological manifestations) can vary substantially
(e.g.,~\cite{Hernandez11}), we first trained individual classifiers
for each participant (rather than a general one for all
participants). The results of our analysis are reported in
Table~\ref{tab:accuracy}. For our analysis, we report values of
accuracy, one of the most commonly used metric to compare performance,
as well as precision and recall of the classes of interest:
'stressed', 'not focused', and 'not awake'. Since the imbalance in the
data can lead to high accuracy values if a classifier always just
predicts the most likely/frequent class while ignoring the class of
higher importance and interest, precision and recall of the class of
interest are also important to
consider~\cite{yap2014,bhattacharyya_data_2011,Hernandez11}.

% However, the imbalance in the data can lead to a classifier with high accuracy if it always just predicts the more likely class while ignoring the class of higher importance and interest in our case (i.e. stressed, not focused, not awake). Therefore, we further report the precision and recall for the three classes 'stressed', 'not focused' and 'not awake'.

Overall, we were able to use extracted physiological features to
predict all three aspects with reasonable accuracy, precision, and
recall, as well as to improve upon the baseline---a stratified random
classifier that randomly chooses one of the two classes with a bias
towards the larger class. While the individually trained classifiers
improved on average across all participants upon the baseline in all
cases excepting recall of 'not focused', the improvement was
substantially higher for awakeness (85\% improvement in precision,
62\% in recall, and 7\% in accuracy) than for stress or focus. Also,
the performance of the individually trained classifiers varied greatly
across participants. While some participants showed a large
improvement, for others the baseline performed much better than the
individually trained classifier. For instance, for predicting
'stressed', the individual classifiers improved upon the baseline for
S4, S6, S8, S11, S12, and S14 with a maximum improvement of 88.4\% in
precision and 184.6\% in recall for S12, while they did worse for S1,
S3, S5, S7, S9, S10, and S13, and in the worst cases did not correctly
predict a single instance of 'stressed'. Typically users that have the
lowest precision and recall values are those where the data is the
most unbalanced. This translates to a scenario where the classifier
will likely label the minority instances as part of the majority class
due to the unbalanced ratio of the data.\\[-0.1cm]
% We attribute these discrepancies to the large differences in response distributions between participants, as well as to the subjectivity of self-reporting.
%
%(improvement: 8\% precision, 1\% recall, 8\% accuracy), the individual performance varied greatly among participants. Some participants showed negligible difference in comparison to the baseline classifier, e.g. subject ..., while others showed a large improvement, e.g. subject ...
%\begin{table}[h!]
%\begin{centering}
%\begin{tabular}{lll}
%\hline
%Participant & Recall Improvement (\%) & Precision Improvement (\%)\\
%\hline
%S12 & 184.6 & 88.4 \\
%S6 & 66.7 & 85.8 \\
%S11 & 50.0 & 44.4 \\
%S8 & 32.4 & 26.7 \\
%S14 & 14.7 & 9.8 \\
%S4 & 5.6 & 5.5 \\
%S9 & -55.4 & -46.8 \\
%S3 & -90.0 & -82.1 \\
%S1 & -100.0 & -100.0 \\
%S5 & -100.0 & -100.0 \\
%S7 & -100.0 & -100.0 \\
%S10 & -100.0 & -100.0 \\
%S13 & -100.0 & -100.0\\
%S2 & - & -\\
%\hline
%\end{tabular}
%\caption{Percentage improvement in recall and precision for stress, using our approach compared to the baseline, on an individual level. For participant 2, the baseline achieved a precision and recall of 0, thus the improvement is undefined}
%\end{centering}
%\label{tab:indImprovement}
%\end{table}

\subsection{Feature Selection and Importance}
%\noindent\textit{Feature Selection and Importance}\\
There are a large variety of features that can be (and have been) calculated in previous research for each of the basic measurements listed in Table~\ref{signals}, such as the mean, standard deviation, maximum, and interquartile range. In addition, each of these metrics can be combined with the various time windows captured of a basic measurement, resulting in a large feature space. To reduce the feature space, we experimented with multiple feature selection methods, including selecting the top k highest correlated features by various metrics such as mutual information, Pearson's correlation coefficient, ANOVA's F-value, as well as wrapper methods such as recursive feature elimination, optimizing mean decrease accuracy by iteratively permuting features, and only selecting features that exceed a certain Gini importance threshold. We found that all methods produced similar results with respect to accuracy, precision, and recall for the individual models. Ultimately, we chose to use the top k features with the highest ANOVA F-value, as it is relatively simple and efficient to calculate. The values of k used were selected by grid search analysis, and are shown in Table \ref{tab:hyperparams}.

%Respiration rate is also highly important for stress (#3 , 14.4%) but I'm not sure what previous works say about correlation with stress
Overall, the features that were selected as the important ones for the individual models based on the random forest algorithm varied greatly across participants. Yet, some feature categories were considered to be important more frequently than others. Table~\ref{tab:featureImportance} shows the averaged Gini importance for the feature categories used for predicting stress. Particularly important for stress were the feature categories heart rate variability (18.3\%) and skin temperature (15.7\%), both of which have been shown in several previous studies to be  indicators for stress~\cite{dishman2000stress,mcduff16,kataoka00}. For predicting focus, the feature categories for blood pulse wave (14.2\%) and heart rate variability (13\%) showed to be the most important categories, while for awakeness the most important ones were heart rate variability (13.6\%) and blood pulse wave (13.1\%).\\[-0.1cm]

\begin{table}[h!]
  \begin{centering}
  \begin{tabular}{llll}
    \hline
    Feature Category & Stress & Focus & Awakeness\\
    \hline
    Heart Rate Variability & 18.3\% & 13\% & 13.6\%\\
    Blood Pulse Wave & 10\% & 14.2\% & 13.1\%\\
    Heart Rate & 8.7\% & 12.6\% & 10.3\%\\
    Skin Temperature & 15.7\% & 9.8\% & 10\%\\
    Galvanic Skin Response & 6.6\% & 8.2\% & 5.1\%\\
    Respiration Rate & 14.8\% & 12.7\% & 10\%\\
    Oxygen Saturation & 5.6\% & 4.3\% & 2\%\\ 
    Energy Expenditure & 6\% & 7.7\% & 4.8\%\\
    Activity & 4.6\% & 7.8\% & 7.6\%\\
    Steps & 1.7\% & 0.8\% & 0.9\%\\
    Time of Day & 0\% & 0.1\% & 0.5\%\\
    \hline
  \end{tabular}
  \caption{The averaged Gini importance of each feature category, per response variable.}
  \label{tab:featureImportance}
  \end{centering}
  \vspace*{-2mm}
\end{table}

\subsection{Individual vs.\ General Model}
%\noindent\textit{Individual vs.\ General Model}\\
Individual models are trained specifically for each individual and thus require a data collection period before they are capable of making accurate predictions. On the other hand, the idea of general models is to be able to train them on already collected data and then  to be able to apply them even to new and unseen individuals, thus overcoming the cold-start problem. Given the large individual differences in biometrics, training a general model to achieve an adequate accuracy for new individuals is not necessarily possible.

To examine the performance of a general model for our participants, we trained three general models, one for focus, one for awakeness and one for stress. We roughly followed the same procedure as for the individual models. Due to the larger amount of data available in the general case, we used the more common random undersampling, which randomly selects elements in the majority class to exclude from the dataset, instead of random oversampling to balance the distribution of the dataset. The models were trained on the datasets of 13 of the 14 participants, and then evaluated on the dataset of the last, repeating this process for all 14 participants.

The bottom row of Table \ref{tab:accuracy} presents the averaged performance results for this approach in terms of accuracy, precision, and recall. Although the averaged precision and recall are comparable or better than those of the averaged individual results, this was at the cost of a large decrease in overall accuracy. Upon closer investigation into the performance of the general model when testing on each participant, we found that individually trained models for each participant performed much better than a general model trained over all participants. Using stress as an example, for participant S12, for whom we saw the greatest increase compared to the baseline in individual models, the general model was unable to predict a single instance of 'stressed' correctly. This is consistent with our expectations because biometric features are highly specific to individuals.

%\todo{put the content of the following sentence somewhere into the discussion; We attribute these discrepancies to the large differences in response distributions between participants, as well as to the subjectivity of self-reporting.}