### 2020-03-26 HICSSSubmission/bp_relatedWork.tex
changed a couple of pics to png because it was messing the document, and slashed a big part of related work, and added spacing so the hicss template works

%!TEX root = bioPrediction_main.tex
\section{Related Work}

The study and prediction approach used in this paper is related
to previous studies of productivity-related indicators and studies
using biometrics to predict them.

\subsection{Productivity-related Indicators}

Our study focuses on the productivity-related indicators of stress, focus and awakeness.

%% Previous approaches~\cite{zuger18,Lalle16,Panwar18} have shown the feasibility of applying machine learning 
%% to biometric data to measure variables such as stress, interruptibility, 
%% negative emotion, or confusion during computer 
%% interaction. These approaches vary in the 
%% variables examined, the invasiveness of the sensors, and the length and 
%% context of the studies. Many previous studies focused on 
%% shorter time frames (mostly hours, but in one case also up to five 
%% days~\cite{sano2013stress} 
%% or two weeks~\cite{zuger18}). To the best of the authors' knowledge, no 
%% other study has examined such an extensive period (eight weeks) in a real-
%% world office environment and the inspected variables (stress, awakeness, and focus).

\vspace{-8mm}

\subsubsection{Stress}

Much previous work relates to identifying and mitigating stress in an office environment.
Previous studies have measured stress by taking one of two possible approaches.
The first approach is to measure plasma catecholamine and cortisol as stress biomarkers~\cite{piazza10}.
This approach is impractical for use over prolonged periods of time, as in our eight-week study.
Further, this approach is imprecise because of the delay from the stress stimulation to the stress response, which 
may take from minutes to hours~\cite{Chandola10,Hellhammer09}.

The second approach, which we have chosen for our study, is to measure autonomic nervous system (ANS) activity by analyzing biometric signals of the human body, such as blood pressure, heartbeat, and temperature~\cite{kataoka00,Eekelen04,valentini10}.
In particular, changes in heart rate variability are associated with cognitive and emotional stress~\cite{mcduff16,dishman2000stress}.
This second approach has been used successfully by several past studies~\cite{Force96,gal07,montano09}.

Hovsepian et al.~\cite{Hovsepian15} have worked to obtain a standard for continuous stress assessment. Their model showed significant improvement over simple heart rate variability measurements.
Hernandez et al.~\cite{Hernandez14} investigated the use of pressure-sensitive keyboard and a capacitative mouse as non-intrusive means for measuring computer users stress levels. They found participants exhibited significantly increased typing pressure and mouse contact when in stressful conditions. McDuff et al.~\cite{mcduff16} experimented with a camera to measure photoplethysmographic signals indicative of cognitive stress. Vizer et al.~\cite{vizer_automated_2009} used keystroke and linguistic features to automatically measure stress levels in response to cognitive and physical stress conditions. All of these studies were performed in a controlled lab setting over a short duration and the results have yet to be replicated in the field.

Evans and Johnson~\cite{evans00} investigated the correlation between noise in the workplace and stress levels. They found that workers exposed to open-office noise showed aftereffects that indicate motivational deficits. Stress in the workplace and its effects on service providers has been analyzed in call centers~\cite{Hernandez11} and in the context of the perceived imbalance between resources and demands~\cite{cherniss80}. Kocielnik et al.~\cite{kocielnik_smart_2013} developed a framework for unobtrusive and continuous measurement of stress in real life conditions. They observed that the data they collected reflected well the participants perceptions of their own stress levels. 
These studies considered several factors such as personality traits, career-related goals and attitudes, and life outside of work, and these factors were correlated with stress levels and burnout.

Our study stands out from these works by nature of its length and focus on knowledge worker stress in everyday office life, using unobtrusive measures. To the best of our knowledge there is no longitudinal study which attempts to explain and predict knowledge worker stress that comes close to the length of our own study. The eight week duration gives us authority to speak on the nature of day-to-day fluctuations in knowledge worker stress levels.

%Bader~\cite{Bader95} proposed an approach towards monitoring and estimating a person's awakeness by leveraging on a stationary pressure sensor to track the person's body movements and a correlator to produce a change signal which is compared with a preset wakefulness threshold in a threshold detector circuit.

%A wearable system for mood assessment considering smartphone features and data from mobile ECGs. Exler, Schankin, 
%Looked at energetic-arousal (tired-awake)
%This paper focuses on valence, and barely mentions energetic arousal 
%so I will leave it out

\vspace{-8mm}

\subsubsection{Focus}

Focus refers to the allocation of limited cognitive processing resources~\cite{Anderson04}.
Mark et al.~\cite{mark2014bored} studied engagement in workplace activities by analyzing the digital activity of 32 information workers in situ for 5 days to understand how attentional states change with context.  They found that doing work that requires focus correlates with stress, while rote work correlates with happiness.

Interruptions in the office are a common barrier keeping workers from sustaining focus including emails, alerts, or interactions with co-workers\cite{gonzalez2004constant,chong2006interruptions,shamsi07}. Interruptions in inopportune times can have negative effects that range from higher error rate and lower overall performance
to an increase in stress and frustration~\cite{bailey2001effects,czerwinski2000instant,mark2008cost}.
External interruptions may cause workers to enter a ``chain of distraction''~\cite{shamsi07}.
This chain is composed by stages of preparation, diversion, resumption and recovery that result in time away from an ongoing task.

Cognitive absorption describes periods of time in which a person experiences total immersion in an activity.
It has been associated with ease of use and perceived usefulness of information technology~\cite{agarwal00}.
Cognitive engagement is described~\cite{webster97} as a period of strong focus in an activity without the feeling of a sense of control of the situation. Flow~\cite{Csikszentmihalyi90}, and mindfulness~\cite{Weick06,dane11} are psychological states that describe periods of prolonged attention and total immersion in an activity.

\vspace{-8mm}

\subsubsection{Awakeness}

Sleepiness (lack of awakeness) has been studied in the context of automobile 
accidents~\cite{connor02,Nordbakke07}.
These studies show a strong association between the level of acute driver sleepiness and the risk of injury crash.
Connor et al.~\cite{connor02} measure sleepiness using the Stanford sleepiness scale, similar to the seven-point Likert scale used in this study.

Nordbakke and Sagberg~\cite{Nordbakke07} show that drivers are well aware of  various factors influencing the risk of falling asleep while driving. Drivers also have good knowledge of the most effective measures to prevent falling asleep at the wheel.
However, most of drivers continue driving even when recognizing sleepiness signals.

\subsection{Biometrics}

Existing work~\cite{sano2013stress,healey2005detecting,wijsman2011towards,zuger2015interruptibility,goyal2017intelligent,Parnin2011,Nakagawa2014,Radevski2015} analyzes a broad array of biometric signals and correlates them with individual's cognitive states and processes.
Zuger et al.~\cite{zuger18} used biometrics to sense interruptibility in the office~\cite{zuger18}.
Biometric signals have also been studied in the context of technology users. Parnin~\cite{Parnin2011} analyzes electromyography to measure sub-vocal utterances, and how these might be correlated with the programmer's perceived difficulty of programming tasks. Similarly, biometrics have been used to measure code difficulty~\cite{fritz2014using}
and using Near Infrared Spectroscopy to measure developer's cerebral blood flow~\cite{Nakagawa2014}.

Eye tracking technology~\cite{Bednarik2006,Crosby1990,Rodeghero2014} and brain activity~\cite{Ikutani2014,Siegmund2014} 
have been used in previous studies to analyze tasks in an office environment.
Eye tracking has been used to analyze memory load and processing load by inspecting task-evoked pupillary response and pupil size~\cite{beatty82}. Similar studies have shown high correlation between pupil size and mental workload of subtasks~\cite{beatty82} and cognitive load~\cite{Klingner10}.
Brain activity has been associated with different mental states~\cite{Berger29} by analyzing specific frequency bands (alpha, beta, gamma, delta, and theta) using electroencephalography (EEG). The increase or decrease of some of these frequencies is correlated with
attentional demand and working memory load~\cite{Smith05,sterman93}.

In contrast to studies that use eye tracking or EEG, we focused on a less invasive technology that can be applied in a real world scenario.
Similarly, we considered adding additional metrics such as Depression, Anxiety and Stress Scale (DASS) and Perceived Stress Scale (PSS)~\cite{Ferdous15} but opted for the ones presented in the paper after several discussions with experts in the area, and after piloting surveys to maximize participant compliance.