### 2020-03-29 IJHCSSubmission/elsarticle-template.tex

%\documentclass[review]{elsarticle}

%\documentclass[preprint,12pt,plainnat]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
   \documentclass[final,1p,times,plainnat]{elsarticle}
%% \documentclass[final,1p,times,twocolumn,authoryear]{elsarticle}
%% \documentclass[final,3p,times,authoryear]{elsarticle}
%% \documentclass[final,3p,times,twocolumn,authoryear]{elsarticle}
%% \documentclass[final,5p,times,authoryear]{elsarticle}
%% \documentclass[final,5p,times,twocolumn,authoryear]{elsarticle}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
%\AtBeginDocument{%
%  \providecommand\BibTeX{{%
%    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

\usepackage{lineno,hyperref}
\modulolinenumbers[5]

\usepackage{balance}       % to better equalize the last page
\usepackage{graphics}      % for EPS, load graphicx instead 
\usepackage[T1]{fontenc}   % for umlauts and other diaeresis
\usepackage{txfonts}
\usepackage{mathptmx}
%\usepackage[pdflang={en-US},pdftex]{hyperref}
\usepackage{color}
\usepackage{booktabs}
\usepackage{textcomp}
\PassOptionsToPackage{hyphens}{url}\usepackage{hyperref}

\renewcommand{\textrightarrow}{$\rightarrow$}

\usepackage{booktabs} % For formal tables
\usepackage{comment}
\usepackage{tabularx}

% Some optional stuff you might like/need.
\usepackage{microtype}        % Improved Tracking and Kerning

\usepackage{todonotes}

\presetkeys{todonotes}{inline}{}
% for commenting and making it visible in the pdf (easy to hide after)
\newboolean{hidecomments}
\setboolean{hidecomments}{false}
%\nochangebars
\ifthenelse{\boolean{hidecomments}}
{\newcommand{\cb}[2]{}}
{\newcommand{\cb}[2]{
		\fbox{\bfseries\sffamily\scriptsize#1}
		{\sf\small$\blacktriangleright$ %$\RHD$
			{#2} $\blacktriangleleft$}}} %\LHD$}}}%
\newcommand\tf[1]{\cb{TF}{\textcolor{green}{#1}}}
\newcommand\cs[1]{\cb{CS}{\textcolor{blue}{#1}}}
\newcommand\ms[1]{\cb{MS}{\textcolor{red}{#1}}}
\newcommand\tfC[1]{\todo[color=green,inline]{Thomas: #1}}

\journal{International Journal of Human-Computer Studies}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frontmatter}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Observing and Predicting Knowledge Worker Stress in the Wild}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Authors will be visible after double blind review}
%\vspace{110cm}
%\vspace{18ex}
%}

\begin{comment}

\author{Mauricio Soto}
\affiliation{%
\institution{ABB Future Labs}
\city{Raleigh}
\state{North Carolina}
}
\email{mauricio.soto@us.abb.com}

\author{Chris Satterfield}
\affiliation{%
\institution{The University of British Columbia}
\city{Vancouver}
\state{British Columbia}
}
\email{cds00@cs.ubc.ca}

\author{Thomas Fritz}
\affiliation{%
\institution{University of Zurich}
\city{Zurich}
\state{Switzerland}
}
\email{fritz@ifi.uzh.ch}

\author{Gail C. Murphy}
\affiliation{%
\institution{The University of British Columbia}
\city{Vancouver}
\state{British Columbia}
}
\email{murphy@cs.ubc.ca}

\author{David Shepherd}
\affiliation{%
\institution{Virginia Commonwealth University}
\city{Richmond}
\state{Virginia}
}
\email{shepherdd@vcu.edu}

\author{Nicholas Kraft}
\affiliation{%
\institution{UserVoice}
\city{Raleigh}
\state{North Carolina}
}
\email{kraft.nicholas.a@gmail.com} \\}

\end{comment}

%% Group authors per affiliation:
%\author{Elsevier\fnref{myfootnote}}
%\address{Radarweg 29, Amsterdam}
%\fntext[myfootnote]{Since 1880.}

%% or include affiliations in footnotes:
%\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
%\ead[url]{www.elsevier.com}

%\author[mysecondaryaddress]{Global Customer Service
%\corref{mycorrespondingauthor}}
%\cortext[mycorrespondingauthor]{Corresponding author}
%\ead{support@elsevier.com}

%\address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
%\address[mysecondaryaddress]{360 Park Avenue South, New York}

\begin{abstract}
Knowledge workers face many challenges in the workplace: work is 
fragmented, disruptions are constant, tasks are complex, and work hours can be 
long. These challenges can impact knowledge workers' stress, focus and awakeness, and in turn their interaction with the digital environment, the quality of work 
performed and their productivity in general. We report on a field study with 14
knowledge workers over an eight-week period in which we investigated, using
experience sampling, how the workers experience stress over time. During this field study, we also collected biometric data, which we then used to investigate if it is possible to predict such productivity-related indicators as stress, focus and awakeness, in the moment.
We observed and report on various trends in knowledge worker stress levels
over several weeks and show that machine learning models can be built to
predict stress, focus and awakeness from biometric data.
\end{abstract}

\begin{keyword}
Biometrics, Stress, Awakeness, Focus, Computer Interaction, Ubiquitous Computing, Empirical Study, User Centered Design
\end{keyword}

\end{frontmatter}

%\linenumbers

%\maketitle

\input{bp_intro_v4}
\input{bp_studyMethod}
%\input{bp_data}
\input{bp_results_1}
%\input{bp_results_2}
\input{bp_discussion}
\input{bp_relatedWork}
\input{bp_threats}
\input{bp_conclusion}

\balance{}

% REFERENCES FORMAT
% References must be the same font size as other body text.
%\bibliographystyle{ACM-Reference-Format}
\bibliography{bp_bibliography}

\end{document}\endinput